---
title: "Project 2"
author: "Rachel Hencher"
date: "2022-09-26"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: pygments
params:
  key: "D3KHf387z9W0EaDoVZNsvD6aOSHPWZmwDvKpTxpr"
  state: "NC"
  ordering: "desc"
  rows: 100
  year: "latest"
  ownership: "private"
  locale: "town"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal for this project is to create a vignette about contacting an API using functions created to query, parse, and return well-structured data. Then, to use the functions to obtain data from the API and do some exploratory data analysis.  

# Required Packages  

`tidyverse`: a collection of R packages that are designed to work together to allow us to read in, transform, and visualize data  
`httr`: allows us to use the `GET` function to access the API   
`jsonlite`: allows us to access the `fromJSON` function to convert JSON data to a data frame  
`ggplot2`: a package in the tidyverse that we will use for creating graphics  
`gridExtra`: allows us to arrange plots in a grid  
`forcats`: TBD
```{r, message=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(ggplot2)
library(gridExtra)
library(forcats)
```

# Functions  
Each of the functions below will contact the chosen API and return well-formatted, parsed data in the form of data frames.  

## `general_info()`  
The following function returns general information on either the largest or smallest *n* colleges in a particular state. The user should provide input for five arguments:  
- `key` requests the user's personal key to access the API  
- `state` requests which state the user would like to retrieve info for using the 2-character abbreviation for the state  
- `size` requests whether the user would like data on large or small schools by selecting "desc" or "asc" respectively  
- `n` is the number of records returned (1-100)  
- `year` requests data from a particular year if the user desires data other than the latest available
```{r}
general_info <- function(key="D3KHf387z9W0EaDoVZNsvD6aOSHPWZmwDvKpTxpr", state="NC", size="desc", n=50, year="latest")
{
  url <- paste0("http://api.data.gov/ed/collegescorecard/v1/schools?api_key=", key, "&school.state=", state, "&per_page=", n, "&sort=student.size:", size, "&_fields=school.name,school.ownership,", year, ".student.size,", year, ".admissions.admission_rate.overall,", year, ".cost.tuition.in_state,", year, ".cost.tuition.out_of_state")
  data <- GET(url)
  parsed_data <- fromJSON(rawToChar(data$content))
  
  parsed_data$results$school.ownership <- as.factor(parsed_data$results$school.ownership)
        levels(parsed_data$results$school.ownership) <- c("Public", "Private, Nonprofit", "Proprietary")
  
  general_data <-
    parsed_data$results %>%
    as_tibble() %>%
    select("Name"=ends_with("name"), "Ownership"=ends_with("ownership"), "Size"=ends_with("size"), "Admissions_Rate"=ends_with("overall"), "In_State_Tuition"=ends_with("in_state"), "Out_of_State_Tuition"=ends_with("out_of_state"))
return(general_data)
}
general_info()
```

## `cost_info()`  
The following function returns cost info on either the most or least expensive *n* colleges in a particular state by in-state tuition. The user should provide input for five arguments:  
- `key` requests the user's personal key to access the API  
- `state` requests which state the user would like to retrieve info for using the 2-character abbreviation for the state  
- `cost` requests whether the user would like data on expensive or more affordable schools by selecting "desc" or "asc" respectively  
- `n` is the number of records returned (1-100)  
- `year` requests data from a particular year if the user desires data other than the latest available
```{r}
cost_info <- function(key="D3KHf387z9W0EaDoVZNsvD6aOSHPWZmwDvKpTxpr", state="NC", cost="desc", n=50, year="latest")
{
  url <- paste0("http://api.data.gov/ed/collegescorecard/v1/schools?api_key=", key, "&school.state=", state, "&per_page=", n, "&sort=cost.tuition.in_state:", cost, "&_fields=school.name,school.ownership,", year, ".cost.tuition.in_state,", year, ".cost.tuition.out_of_state,", year, ".cost.booksupply,", year, ".cost.roomboard.oncampus,", year, ".cost.roomboard.offcampus")
  data <- GET(url)
  parsed_data <- fromJSON(rawToChar(data$content))
  cost_data <-
    parsed_data$results %>%
    as_tibble() %>%
    select("Name"=ends_with("name"), "Ownership"=ends_with("ownership"), "In_State_Tuition"=contains("in"), "Out_of_State_Tuition"=contains("out"), "Books_Supplies"=ends_with("supply"), "Room_Board_On"=ends_with("oncampus"), "Room_Board_Off"=ends_with("offcampus"))
  
    cost_data$Ownership <- as.factor(cost_data$Ownership)
      levels(cost_data$Ownership) <- c("Public", "Private, Nonprofit", "Proprietary")
return(cost_data)
}
cost_info()
```

## `admissions_info()`  
The following function returns admissions info on *n* colleges under a particular "ownership" category. The user should provide input for five arguments:  
- `key` requests the user's personal key to access the API  
- `ownership` requests whether the user would like to retrieve admissions info for "public", "private" (nonprofit), or "proprietary" colleges  
- `rate` requests whether the user would like data on less competitive or more selective schools by selecting "desc" or "asc" respectively  
- `n` is the number of records returned (1-100)  
- `year` requests data from a particular year if the user desires data other than the latest available
```{r}
admissions_info <- function(key="D3KHf387z9W0EaDoVZNsvD6aOSHPWZmwDvKpTxpr", ownership="private", rate="asc", n=50, year="latest")
{
  ownership <- ifelse(tolower(ownership)=="public", "1", 
                      ifelse(tolower(ownership)=="private", "2", 
                             ifelse(tolower(ownership)=="proprietary", "3", "ERROR"))) 
        
  url <- paste0("http://api.data.gov/ed/collegescorecard/v1/schools?api_key=", key, "&school.ownership=", ownership, "&per_page=", n, "&sort=admissions.admission_rate.overall:", rate, "&school.degrees_awarded.predominant=3&_fields=school.name,", year, ".admissions.admission_rate.overall,", year, ".admissions.test_requirements,", year, ".admissions.sat_scores.midpoint.critical_reading,", year, ".admissions.sat_scores.midpoint.writing,", year, ".admissions.sat_scores.midpoint.math,", year, ".admissions.act_scores.midpoint.english,", year, ".admissions.act_scores.midpoint.writing,", year, ".admissions.act_scores.midpoint.math")
  data <- GET(url)
  parsed_data <- fromJSON(rawToChar(data$content))
  
  admissions_data <-
    parsed_data$results %>%
    as_tibble() %>%
    select("Name"=ends_with("name"), "Ownership"=ends_with("ownership"), "Admissions_Rate"=ends_with("overall"), "Test_Requirements"=ends_with("requirements"), "SAT_Reading"=ends_with("reading"), "SAT_Writing"=ends_with("sat_scores.midpoint.writing"), "SAT_Math"=ends_with("sat_scores.midpoint.math"), "ACT_English"=ends_with("english"), "ACT_Writing"=ends_with("act_scores.midpoint.writing"), "ACT_Math"=ends_with("act_scores.midpoint.math"))
    
    admissions_data$Test_Requirements <- as.factor(admissions_data$Test_Requirements)
      levels(admissions_data$Test_Requirements) <- c("Required", "Recommended", "Neither required nor recommended", "Do not know", "Considered but not required")
  
return(admissions_data)
}
admissions_info()
```

## `demographic_info()`  
The following function returns demographic  information of the school based on the location info on *n* colleges. The user should provide input four arguments:  
- `key` requests the user's personal key to access the API  
- `locale` requests the user to mention if they would like to see data on the location of the school.(city/suburb/town/rural)
- `rate` requests whether the user would like data on the number of students in the school by selecting "desc" or "asc" respectively  
- `n` is the number of records returned (1-100)

```{r}
demographic_info <- function(key="k42psBgICW3DEe6eS1gc7AxbTTYtOwOGN9URGVuT", state = "NC", size = "asc", locale="town", n=50, year= "latest")
{
  location <- ifelse(locale=="city", "11,12,13", 
                     ifelse(locale=="suburb", "21,22,23",
                            ifelse(locale=="town", "31,32,33",
                                   ifelse(locale=="rural", "41,42,43", "ERROR"))))
  
    url <- paste0("http://api.data.gov/ed/collegescorecard/v1/schools?api_key=", key, "&school.state=", state, "&school.locale=", location, "&per_page=", n, "&sort=student.size:", size, "&_fields=school.name,school.ownership,school.city,", year, ".student.size,", year, ".student.demographics.race_ethnicity.aian,", year, ".student.demographics.race_ethnicity.nhpi,", year, ".student.demographics.race_ethnicity.asian,", year, ".student.demographics.race_ethnicity.black,", year, ".student.demographics.race_ethnicity.white,", year, ".student.demographics.race_ethnicity.hispanic,", year, ".student.demographics.race_ethnicity.unknown,", year, ".student.demographics.men,", year, ".student.demographics.women")
  
  data <- GET(url)
  parsed_data <- fromJSON(rawToChar(data$content))
  
  demographic_data <-
    parsed_data$results %>%
    as_tibble() %>%
    select("Name"=ends_with("name"),
           "City"=ends_with("city"),
           "Ownership"=ends_with("ownership"),
           "Total_Enrollment"=ends_with("size"),
           "AIAN"=ends_with("aian"),
           "NHPI"=ends_with("nhpi"),
           "Asian"=ends_with("asian"),
           "Black"=ends_with("black"),
           "White"=ends_with("white"),
           "Hispanic"=ends_with("hispanic"),
           "Unknown"=ends_with("unknown"),
           "Men"=ends_with("men"),
           "Women"=ends_with("women")
           )
  
  demographic_data$Ownership <- as.factor(demographic_data$Ownership)
        levels(demographic_data$Ownership) <- c("Public", "Private, Nonprofit", "Proprietary")
  
  demographic_data <- demographic_data %>% 
    mutate(Total_AIAN=round(Total_Enrollment*AIAN), 
           Total_NHPI=round(Total_Enrollment*NHPI), 
           Total_Asian=round(Total_Enrollment*Asian), 
           Total_Black=round(Total_Enrollment*Black),
           Total_White=round(Total_Enrollment*White), 
           Total_Hispanic=round(Total_Enrollment*Hispanic),
           Total_Unknown=round(Total_Enrollment*Unknown),
           Total_Men=round(Total_Enrollment*Men1),
           Total_Women=round(Total_Enrollment*Women)) %>% 
    select(Name, City, Ownership, starts_with("Total_")) %>% 
    pivot_longer(cols=!c(Name, City, Ownership), names_to="Ethnicity_Gender", values_to="Count") 
  
return(demographic_data)
}
demographic_info()
```

## `financial_info()`  
The following function returns student financial information on either the largest or smallest *n* colleges in a particular state. The user should provide input for five arguments:  
- `key` requests the user's personal key to access the API  
- `state` requests which state the user would like to retrieve info for using the 2-character abbreviation for the state  
- `size` requests whether the user would like data on large or small schools by selecting "desc" or "asc" respectively  
- `n` is the number of records returned (1-100)  
- `year` requests data from a particular year if the user desires data other than the latest available

```{r}
financial_info <- function(key="D3KHf387z9W0EaDoVZNsvD6aOSHPWZmwDvKpTxpr", state="NC", size="desc", n=50, year="latest")
{
  url <- paste0("http://api.data.gov/ed/collegescorecard/v1/schools?api_key=", key, "&school.state=", state, "&per_page=", n, "&sort=student.size:", size, "&_fields=school.name,school.ownership,school.city,", year, ".student.size,", year, ".student.demographics.poverty_rate,", year, ".student.demographics.median_family_income,", year, ".aid.median_debt.female_students,", year, ".aid.median_debt.male_students")
  data <- GET(url)
  parsed_data <- fromJSON(rawToChar(data$content))
  
  financial_data <- 
    parsed_data$results %>%
    as_tibble() %>%
    select("Name"=ends_with("name"), "Ownership"=ends_with("ownership"), "City"=ends_with("city"), "Size"=ends_with("size"), "Poverty_Rate"=ends_with("poverty_rate"), "Median_Family_Income"=ends_with("income"), "Female_Median_Debt"=ends_with("female_students"), "Male_Median_Debt"=ends_with("debt.male_students"))
  
    financial_data$Ownership <- as.factor(financial_data$Ownership)
      levels(financial_data$Ownership) <- c("Public", "Private, Nonprofit", "Proprietary")
return(financial_data)
}
financial_info()
```

## `earnings_info()`  
The following function returns earnings information on *n* colleges with either the highest or lowest median student earnings 10 years after entry. The user should provide input for four arguments:  
- `key` requests the user's personal key to access the API  
- `earnings` requests whether the user would like data on schools whose students are the highest earners or lowest earners by selecting "desc" or "asc" respectively  
- `n` is the number of records returned (1-100)  
- `year` requests data from a particular year if the user desires data other than the latest available

```{r}
earnings_info <- function(key="D3KHf387z9W0EaDoVZNsvD6aOSHPWZmwDvKpTxpr", earnings="desc", n=50, year="latest")
{
  url <- paste0("http://api.data.gov/ed/collegescorecard/v1/schools?api_key=", key, "&per_page=", n, "&sort=earnings.10_yrs_after_entry.median:", earnings, "&school.degrees_awarded.predominant=3&_fields=school.name,", year, ".cost.tuition.in_state,", year, ".cost.tuition.out_of_state,", year, ".earnings.10_yrs_after_entry.median,", year, ".earnings.10_yrs_after_entry.mean_earnings.female_students,", year, ".earnings.10_yrs_after_entry.mean_earnings.male_students")
  data <- GET(url)
  parsed_data <- fromJSON(rawToChar(data$content))
  earnings_data <-
    parsed_data$results %>%
    as_tibble() %>%
    select("Name"=ends_with("name"), "In_State_Tuition"=ends_with("in_state"), "Out_of_State_Tuition"=ends_with("out_of_state"), "Median_Earnings"=ends_with("median"), "Mean_For_Females"=ends_with("female_students"), "Mean_For_Males"=ends_with("mean_earnings.male_students"))
return(earnings_data)
}
earnings_info()
```

# Data Exploration
```{r, warning=FALSE, include=TRUE, echo=TRUE}
exploration <- function(general_data, cost_data, admissions_data, demographic_data, financial_data, earnings_data, state, locale) {
summary(general_data)
  
plot1 <- ggplot(general_data, aes(x=Ownership, y=Size)) + 
  geom_boxplot(fill = "#112446") + 
  labs(title="Size by Ownership Type for State Colleges") + 
  theme_minimal()
plot1
plot2 <- ggplot(general_data, aes(x=Admissions_Rate, y=In_State_Tuition)) + 
  geom_point(aes(color=Ownership)) + 
  labs(x="Admissions Rate", y="In-State Tuition", title="In-State Tuition vs Admissions Rate for State Colleges") + 
  theme_minimal()
plot2
plot3 <- ggplot(general_data, aes(x=Admissions_Rate, y=Out_of_State_Tuition)) + 
  geom_point(aes(color=Ownership)) + 
  labs(x="Admissions Rate", y="Out-of-State Tuition", title="Out-of-State Tuition vs Admissions Rate for State Colleges") + 
  theme_minimal()
plot3
admissions_data2 <- admissions_data %>%
  mutate("Total_SAT"=(sumrow=SAT_Reading+SAT_Math), "Total_ACT"=(sumrow=ACT_English+ACT_Math))
plot4 <- ggplot(admissions_data2, aes(x=Admissions_Rate, y=Total_SAT)) + 
  geom_point(aes(shape=Test_Requirements, color=Test_Requirements)) + 
  geom_smooth(method=lm) + labs(x="Admissions Rate", y="Total SAT for Reading/Math", title="Total SAT Score vs Admissions Rate for US Colleges") + 
  theme_minimal()
plot4
plot5 <- ggplot(admissions_data2, aes(x=Admissions_Rate, y=Total_ACT)) + 
  geom_point(aes(shape=Test_Requirements, color=Test_Requirements)) + 
  geom_smooth(method=lm) + 
  labs(x="Admissions Rate", y="Total ACT for English/Math", title="Total ACT Score vs Admissions Rate for US Colleges") + 
  theme_minimal()
plot5
plot6 <- ggplot(admissions_data2, aes(x=Total_SAT, y=Total_ACT)) + 
  geom_point(aes(shape=Test_Requirements, color=Test_Requirements)) + 
  geom_smooth(method=lm) + 
  labs(x="Total SAT for Reading/Math", y="Total ACT for English/Math", title="Total ACT Score vs Total SAT Score for US Colleges") + 
  theme_minimal()
plot6
plot7 <- ggplot(cost_data, aes(x=Ownership, y=In_State_Tuition)) + 
  geom_boxplot(fill="#112446") + 
  scale_y_continuous(trans="log10") + 
  geom_jitter() +
  labs(y="In-State Tuition", title="In-State Tuition by Ownership Type for State Colleges") + 
  theme_minimal()
plot7
plot8 <- ggplot(cost_data, aes(x=Ownership, y=Out_of_State_Tuition)) + 
  geom_boxplot(fill="#112446") + 
  scale_y_continuous(trans="log10") + 
  geom_jitter() +
  labs(y="Out-of-State Tuition", title="Out-of-State Tuition by Ownership Type for State Colleges") + 
  theme_minimal()
plot8
demo <- ggplot(demographic_data, aes(Ownership, Count)) +
  geom_col(fill = "#112446") +
  labs(
    x="Ownership ",
    y="Total Enrollment",
    title="Undergraduate Students enrolled for each Ownership",
    subtitle=paste("State :", state, "; Locale :", locale)
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) 
demographic_data_ownership <- demographic_data %>% 
  filter(!(Ethnicity_Gender %in% c("Total_Men", "Total_Women")), !(Ownership %in% "Proprietary")) %>% 
  group_by(Ownership, Ethnicity_Gender) %>% 
  summarise(Total_Count=sum(Count)) %>% 
  arrange(Total_Count, .by_group = TRUE)
demographic_data_ownership$Ethnicity_Gender < factor(demographic_data_ownership$Ethnicity_Gender, levels=demographic_data_ownership[order(demographic_data_ownership$Ethnicity_Gender), "Total_Count"])
demographic_data_ownership
demographic_o <- demographic_data_ownership %>% 
  filter(!(Ethnicity_Gender %in% "Total_Enrollment")) %>%
  arrange(Ethnicity_Gender) %>% 
  ggplot(aes(x=Ownership, y=Total_Count, fill=Ethnicity_Gender)) +
  geom_bar(position=position_dodge2(reverse=TRUE), stat='identity') +
  scale_fill_hue(direction=1) +
  labs(
    x="Ownership",
    y="Total Student Count",
    title="Undergraduate Students enrolled for each Ownership by Ethnicity",
    subtitle=paste("State :", state, "; Locale :", locale)
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) 
demographic_data_gender <- demographic_data %>% 
  filter(Ethnicity_Gender %in% c("Total_Men", "Total_Women")) %>% 
  ggplot(aes(x=Ownership, y=Count, fill=Ethnicity_Gender)) +
  geom_boxplot() +
  scale_y_continuous(trans="log10") +
  labs(
    x="Ownership",
    y="Count",
    title="Count of Students for each Ownership by Gender",
    subtitle=paste("State :", state, "; Locale :", locale)
  ) + 
  theme_minimal() +
  theme(plot.title=element_text(hjust=0.5),
        plot.subtitle=element_text(hjust=0.5)) 
  
financial_data_pr <- financial_data %>% 
  group_by(Ownership, City) %>% 
  summarise(Median_Poverty_Rate=median(Poverty_Rate, na.rm=TRUE))
financial_data_pr
financial_pr <- financial_data_pr %>%
  filter(!is.na(Median_Poverty_Rate)) %>%
  ggplot() +
  aes(x=Median_Poverty_Rate, y=..density..) +
  geom_histogram(bins=14L, fill="#112446") +
  geom_density(color="red", 
               size=2) +
  labs(
    x="Poverty Rate",
    title="Poverty Rate Distribution by State"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(hjust = 0.5),
        plot.subtitle=element_text(hjust = 0.5)) 
poverty_ownership <- financial_data %>% 
  group_by(Ownership) %>% 
  summarise(Mean_PR = mean(Poverty_Rate, na.rm=TRUE)) %>% 
  filter(!(Ownership %in% "Proprietary"))
poverty_o <- ggplot(poverty_ownership) +
  aes(x=Ownership, y=Mean_PR) +
  geom_col(fill="#112446") +
  labs(
    x="Poverty Rate",
    title="Poverty Rate by Ownership for Each State"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(hjust = 0.5),
        plot.subtitle=element_text(hjust = 0.5)) 
instate_earning <- ggplot(earnings_data) +
  aes(x=In_State_Tuition, y=Median_Earnings) +
  geom_point(shape="circle", size=2, colour="#112446") +
  geom_smooth(span=0.75) +
  theme_minimal() +
  theme(plot.title=element_text(hjust=0.5),
        plot.subtitle=element_text(hjust=0.5)) 
outstate_earning <-  ggplot(earnings_data) +
  aes(x=Out_of_State_Tuition, y=Median_Earnings) +
  geom_point(shape="circle", size=2, colour="#112446") +
  geom_smooth(span=0.75) +
  theme_minimal() +
  theme(plot.title=element_text(hjust=0.5),
        plot.subtitle=element_text(hjust=0.5)) 
earnings <- grid.arrange(instate_earning, outstate_earning, ncol=2)
return(list(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, demo, demographic_o, demographic_data_gender, financial_pr, poverty_o, earnings))
}
```


# Main Function 
Take the user input of:
- state
- ordering 
- rows 
- year 
- ownership 
- locale


```{r, warning=FALSE, include=TRUE, echo=TRUE}
main <- function() {
  key = params$key
  state = params$state
  size = params$ordering
  n = params$rows 
  year = params$year
  ownership = params$ownership
  locale = params$locale
  general_data <- general_info(key, state, size, n, year)
  cost_data <- cost_info(key, state, cost = size, n, year)
  admissions_data <- admissions_info(key, ownership, rate=size, n, year)
  demographic_data <- demographic_info(key, state, size, locale, n, year)
  financial_data <- financial_info(key, state, size, year)
  earnings_data <- earnings_info(key, earnings = size, n, year)
  exploration(general_data, cost_data, admissions_data, demographic_data, financial_data, earnings_data, state, locale)
}
```

```{r, warning=FALSE, include=TRUE, echo=TRUE}
main()
```
